---
title: "Lab 03: Alumni Jobs + Salaries"
subtitle: "due Sun, Feb 14 at 11:59p EST"
output: 
  tufte::tufte_html:
    css: "./sta210-labs.css"
    tufte_variant: "envisioned"
    highlight: pygments
    fig_height: 10
    toc: true
    toc_depth: 1
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning=FALSE, 
                      message=FALSE)
```

The primary goal of today's lab is to practice with some of the primary data wrangling, data visualization, and modeling functions we will use for regression analysis. It is also another opportunity to continue practicing using RStudio and GitHub before you start collaborating with others.

# Learning objectives

-   Fitting a simple linear regression model using the `lm` function
-   Interpreting the slope and intercept
-   Using statistical inference to draw conclusions about the slope

Meet your team!

See STA 210 teams to see your team assignment. This will be your team for labs and the final project.

Before you get started on the lab assignment, we will take a few minutes to help you develop a plan for working as a team.

`r emo::ji("white_check_mark")` Come up with a team name. I encourage you to be creative! Your TA will get your team name by the end of lab.

`r emo::ji("white_check_mark")` Identify something everyone on the team has in common that's not necessarily in common with everyone else in the class.

`r emo::ji("white_check_mark")` Fill out the team agreement. This will help you figure out a plan for working together during labs and outside of lab times. You can find the team agreement in the GitHub repo **team-agreement-[github_team_name]**.

-   Have **one person** from the team clone the repo and start a new RStudio project. This person will type the team's responses as you discuss the sections of the agreement. Share your screen, if possible, so the rest of the team can see the updates. No one else in the team should type at this point but should be contributing to the discussion.
-   Be sure to push the completed agreement to GitHub. Each team member can refer to the document in this repo or download the PDF of the agreement for future reference.

# Getting Started

A repository has already been created for you and your teammates. Everyone in your team has access to the same repo.

-   Go to [STA 210 course organization](https://www.github.com/sta210-sp21) on GitHub.

-   In addition to your private individual repositories, you should now see a repo named **lab-03**. Go to that repository.

-   Each person on the team should clone the repository and open a new project in RStudio. **Do not make any changes to the .Rmd file until the instructions tell you do to so.**

## Workflow: Using git and GitHub as a team

**Assign each person on your team a number 1 through 4. For teams of three, Member 1 can take on the role of Member 4.**

The following exercises must be done in order. **Only one person should type in the .Rmd file and push updates at a time.** When it is not your turn to type, you should still share ideas and contribute to the team's discussion.

## Update YAML

**Team Member 1: Change the author to your team name and include each team member's name in the `author` field of the YAML in the following format. `Team Name: Member 1, Member 2, Member 3, Member 4`.**

# Packages

We will use the following packages in today's lab.

```{r load-packages}
library(tidyverse)
library(knitr)
library(broom)
```

# Data: Recent college graduates

Today's data are two data sets that were part of [TidyTuesday College tuition, diversity, and pay](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-10/readme.md). All of the information is originally from the Department of Education but was collected from various websites.

We will use the following two data sets in this lab:

## `alumni` data set

The information in this data set was collected from the [PayScale College Salary Report](https://www.payscale.com/college-salary-report/bachelors).

+--------------------------------------+--------------+----------------------------------------------------------------------------------+
| variable                             | class        | description                                                                      |
+:=====================================+:=============+:=================================================================================+
| rank                                 | double       | Potential salary rank within state                                               |
+--------------------------------------+--------------+----------------------------------------------------------------------------------+
| name                                 | character    | Name of school                                                                   |
+--------------------------------------+--------------+----------------------------------------------------------------------------------+
| state_name                           | character    | state name                                                                       |
+--------------------------------------+--------------+----------------------------------------------------------------------------------+
| early_career_pay                     | double       | Median salary for alumni with 0 - 5 years experience (in US dollars)             |
+--------------------------------------+--------------+----------------------------------------------------------------------------------+
| mid_career_pay                       | double       | Median salary for alumni with 0 - 5 years experience (in US dollars)             |
+--------------------------------------+--------------+----------------------------------------------------------------------------------+
| make_world_better_percent            | double       | Percent of alumni who think they are making the world a better place             |
+--------------------------------------+--------------+----------------------------------------------------------------------------------+
| stem_percent                         | double       | Percent of degrees awarded in science, technology, engineering, or math subjects |
+--------------------------------------+--------------+----------------------------------------------------------------------------------+

## `tuition` data set

The information in this data set was collected from [Tuition Tracker](https://www.tuitiontracker.org/).

+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| variable                         | class         | description                                                                           |
+:=================================+:==============+:======================================================================================+
| name                             | character     | School name                                                                           |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| state                            | character     | State name                                                                            |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| state_code                       | character     | State Abbreviation                                                                    |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| type                             | character     | Type: Public, private, for-profit                                                     |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| degree_length                    | character     | 4 year or 2 year degree                                                               |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| room_and_board                   | double        | Room and board in USD                                                                 |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| in_state_tuition                 | double        | Tuition for in-state residents in USD                                                 |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| in_state_total                   | double        | Total cost for in-state residents in USD (sum of room & board + in state tuition)     |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| out_of_state_tuition             | double        | Tuition for out-of-state residents in USD                                             |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+
| out_of_state_total               | double        | Total cost for in-state residents in USD (sum of room & board + out of state tuition) |
+----------------------------------+---------------+---------------------------------------------------------------------------------------+

## Load data

The data set for this lab will focus on higher education institutions where the median early career pay is \$70,000 or less and that are included in both data sets.


```{r echo = F, eval = F}
#clean data to prepare for lab
# remove duplicates + only keep observations in both data sets
alumni <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

alumni_keep <- semi_join(alumni, tuition_cost, by = c("name", "state_name" = "state"))

alumni_keep %>% write_csv("data/alumni-salaries.csv")

tuition_keep <- semi_join(tuition_cost, alumni,  c("name", "state" = "state_name"))

tuition_keep %>% write_csv("data/tuition.csv")
```

```{r}
alumni <- read_csv("data/alumni-salaries.csv")
tuition <- read_csv("data/tuition.csv")
```

# Exercises

## Instructions

Is there a relationship between someone's early career salary and whether they perceive their job as making the world a better place? To answer this question, we will use linear regression to fit a model using the percent of alumni who perceive their job is making the world a better place to predict the median early career salary for colleges and universities in the United States.

::: {.commit}
<b>Team Member 1</b>: Type the team's responses to exercises 1 - 4.
:::

1.  Plot the distribution of the response variable and calculate the appropriate summary statistics. Describe the distribution including the shape, center, spread, and any outliers.

2.  Plot the distribution of explanatory variable and calculate the appropriate summary statistics. Describe the distribution including the shape, center, spread, and any outliers.

3.  What do you expect the relationship to be between the percent of alumni who perceive their jobs of making the world a better place and the typical career salary? Discuss as a group and summarize the group's thoughts.

4.  Create a visualization to display the relationship between `early_career_pay` and `make_world_better_percent`. Use the scatter plot to describe the relationship between the two variables. Is the relationship what your group expected? Why or why not?

::: {.commit}
`r emo::ji("white_check_mark")` `r emo::ji("arrow_up")` <b>Team Member 1</b>: Knit, commit and push your changes to GitHub with an appropriate commit message again. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.

</b>All other team members</b>: <b>Pull</b> to get the updated documents from GitHub. Click on the .Rmd file, and you should see the responses to exercises 1- 4.

<b>Team Member 2</b>: It's your turn! Type the team's response to exercises 5 - 7.
:::

1.  Fit a simple linear regression model for the relationship between early career pay and perceiving their job as making the world a better place. Display the code and output, include a confidence interval for the coefficient, and use the `kable` function to neatly display the model output with 3 digits.

2.  Interpret the slope in the context of the problem.

3.  Does it make sense to interpret the intercept? If so, interpret the intercept. Otherwise, briefly explain why not.

::: {.commit}
`r emo::ji("white_check_mark")` `r emo::ji("arrow_up")` <b>Team Member 2</b>: Knit, commit and push your changes to GitHub with an appropriate commit message again. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.

</b>All other team members</b>: <b>Pull</b> to get the updated documents from GitHub. Click on the .Rmd file, and you should see the responses to exercises 5 - 7.

<b>Team Member 3</b>: It's your turn! Type the team's response to exercises 8 - 10.
:::

1.  Is the linear relationship between `early_career_pay` and `make_world_better_percent` statistically significant?

    -   State the null and alternative hypotheses used to answer this question in words and in mathematical notation.
    -   What is the test statistic? Write the equation used to calculate the test statistic.
    -   What distribution was used to calculate the p-value?
    -   State your conclusion for the test in the context of the data.

2.  Interpret the 95% confidence interval for the slope in the context of the data.

```{marginfigure}
See [Section 13.4 Mutating joins](https://r4ds.had.co.nz/relational-data.html#mutating-joins) in R for Data Science for more details about joins.
```

1.  For the remainder of the lab, we will consider differences between private and public institutions. The institution type is in the `tuition_cost` data frame. Join `tuition_cost` and `alumni` to create a single data frame. *Hint: You need to consider the name and state, since there are some states with colleges of the same name.*

::: {.commit}
`r emo::ji("white_check_mark")` `r emo::ji("arrow_up")` <b>Team Member 3</b>: Knit, commit and push your changes to GitHub with an appropriate commit message again. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.

</b>All other team members</b>: <b>Pull</b> to get the updated documents from GitHub. Click on the .Rmd file, and you should see the responses to exercises 8 - 10. </b>

<b>Team Member 4</b>: It's your turn! Type the team's response to exercises 11 - 13.
:::

1.  Fit a model for the relationship between `early_career_pay` and `make_world_better_percent` for **Private** institutions. Use the `kable` function to neatly display the model output with 3 digits and confidence intervals for the slope and intercept.

    -   Interpret the interval in the context of the data.

2.  Fit a model for the relationship between `early_career_pay` and `make_world_better_percent` for **Public** institutions. Use the `kable` function to neatly display the model output with 3 digits and confidence intervals for the slope and intercept.

    -   Interpret the interval in the context of the data.

3.  Based on the previous exercises, is there evidence that does the relationship between `early_career_pay` and `make_world_better_percent` differs significantly by institution type? Briefly explain your reasoning.

<div class = "commit">

`r emo::ji("white_check_mark")` `r emo::ji("arrow_up")` <b>Team Member 4</b>: Knit, commit and push your changes to GitHub with an appropriate commit message again. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.

<b>All other team members</b>: <b>Pull</b> to get the updated documents from GitHub. Click on the .Rmd file, and you should see the team's completed lab! </b>

# Wrapping up

Go back through your write up to make sure you followed the coding style guidelines we discussed in class (e.g. no long lines of code)

<b>Team Member 2:</b> Make any edits as needed. Then knit, commit, and push the updated documents to GitHub if you made any changes.

All other team members can click to pull the finalized document.

# Submission

<b>Team Member 3</b>: Upload the team's PDF to Gradescope. Be sure to include every team member's name in the Gradescope submission Associate the "Overall" graded section with the first page of your PDF, and mark where each answer is to the exercises. If any answer spans multiple pages, then mark all pages.

There should only be <u>one</u> submission per team on Gradescope.

<br>
