---
title: "Project topic ideas" 
subtitle: "due Sun, Mar 07, at 11:59p EST"
output: 
  tufte::tufte_html:
    css: "./sta210-labs.css"
    tufte_variant: "envisioned"
    highlight: pygments
    fig_height: 10
    toc: true
    toc_depth: 1
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

In this lab you will find 2 - 3 potential data sets to use for the final project. The goal is to give you enough time to find a data set that will help you be successful on the project. 

# Getting started

## Clone assignment repo + start new project

A repository has already been created for you and your teammates. Everyone in your team has access to the same repo.

- Go to [course organization](https://www.github.com/sta210-sp21) on GitHub.

- In addition to your private individual repositories, you should now see a repo named **project-**. Go to that repository.
  
- Each person on the team should clone the repository and open a new project in RStudio.

# Topic ideas 

- Go to the [Project](https://sta210-sp21.netlify.app/project/) page on the course website and read in the introduction and **Topic Ideas** sections.

- As you identify data sets for the project, make sure they meet the criteria stated under the **Topic Ideas** section.

- See the [Tips + Resources](https://sta210-sp21.netlify.app/project/#tips--resources) section of the project instructions for a list of data repositories to help you find potential data sets.

**Note**: The `topic-ideas.Rmd` file and data sets are stored in different folders, so you will need to use `../data/` before the file name to read in the data set. 

For example, 

```{r eval = F}
my_data_frame <- read_csv("../data/my_data.csv")
```

# Submission

You will submit the final project ideas by pushing your work to your GitHub repo. The teaching team will provide feedback under the **Issues** on GitHub. 

You do **<u>not</u>** need to submit anything for the topic ideas on Gradescope. 